Arc Blanc: a real time ocean simulation framework

### 2025-01-10 paper.tex
Use yuksel 2007 in reference
Change tessendorf references

While there are many articles describing different parts of ocean simulation, the literature lacks papers that bring all the pieces together.
This article proposes a fully described real-time framework designed to simulate the free surface of the ocean and the interactions between solid objects and the surrounding fluid in real time (see Figure \ref{fig:overviewFramework}).
This new framework incorporates three different methods from the literature:
\begin{itemize}
	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
	\item The estimation of the forces applied by the fluid on a solid mesh, mixing together different techniques \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014}.
	\item The simulation of the alteration of a free surface by a solid, inspired by Cords and Staadt works (2009) \cite{cordsRealTimeOpenWater2009}.
\end{itemize}
To bring together these previous works, this new framework proposes two new contributions:
\begin{itemize}
	\item An algorithm to compute the ocean velocity at any depth from the simulation of the free surface. This velocity is applied to the fluid to solid forces.
	\item An algorithm to compute the input of the Cords and Staadt method more realistically from the free surface.
\end{itemize}

@@

\subsection{Ocean free surface simulation}
\label{subsubsec:oceanFreeSurfaceSim}
J. Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} presented what could be considered a seminal work in real-time ocean simulation. His work relies on the Fast Fourier Transform (FFT) and statistical wave spectra to accurately find the analytical solution to an approximation of the Navier-Stokes equations on a 2D height field. This method is commonly known as the “Tessendorf method”.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solid to fluid coupling}
\label{subsubsec:solidToFluidCoupling}
The simulation of the surface of water in contact with a solid body has been an active research topic for the last twenty years.
Gomez (2000) \cite{gomezInteractiveSimulationWater2000} solved the wave equation using the finite-difference method (FDM) to simulate waves on the free surface.
This method was improved by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to work with moving objects.
Tessendorf (2004) \cite{tessendorfInteractiveWaterSurface2004} proposed a convolutional method.
He extended his method by using an exponential solver \cite{tessendorfEWaveUsingExponential2014}.
Yuksel (2010) \cite{yukselRealtimeWaterWaves2010} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
Chentanez and Müller (2010) \cite{chentanezRealTimeSimulationLarge2010} created a hybrid solver combining Eulerian and Lagrangian approaches.
Canabal  \etal (2016) \cite{canabalDispersionKernelsWater2016} improved the physical accuracy of the convolutional method by introducing a dispersion kernel.
Jeschke and Woktan (2017) \cite{jeschkeWaterWavePackets2017} took the work of C. Yuskel a bit further by introducing the notion of a wave packet, which carries information about an entire wave train.
One year later, Jeschke  \etal (2018) \cite{jeschkeWaterSurfaceWavelets2018} proposed to use a wavelet transformation that discretizes the wave amplitudes as a function of space, frequency and direction.
Finally, Schreck  \etal (2019) \cite{schreckFundamentalSolutionsWater2019} based their work on the method of fundamental solutions to generate ambient waves on a large scale domain interacting with complex boundaries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Water height}
\label{subsec:waterHeight}
Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
\begin{equation}
	\label{equa:heightWater}
	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(\mathbf{k}, t) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
\end{equation}
where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
\begin{equation}
	\label{equa:heightWaterBis}
	\tilde{h}(\mathbf{k}, t) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
\end{equation}
where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see Appendix~\ref{sec:annexe} for more details.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many different \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
In its original paper, Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
\begin{equation}
	\tilde{h}_0(\mathbf{k}) = \xi_{\mathbf{k}}\sqrt{\frac{4\pi}{L_ik}S(\omega)D(\omega, \theta)\left|\frac{\partial\omega(k)}{\partial k}\right|}
\end{equation}
where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphes \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.

@@

\Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x\ D_z\right]^T$ of seawater to simulate more realistic choppy waves.
As proposed in Sections 4.5 and 4.6 of Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999}, it is calculated as follows:
\begin{equation}\label{equa:hori-displacement}
	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \frac{i\mathbf{k}}{k} \tilde{h}(\mathbf{k}, t)\exp\left(i\mathbf{k}\cdot\mathbf x\right).
\end{equation}

>>>>

While there are many articles describing different parts of ocean simulation, the literature lacks papers that bring all the pieces together.
This article proposes a fully described real-time framework designed to simulate the free surface of the ocean and the interactions between solid objects and the surrounding fluid in real time (see Figure \ref{fig:overviewFramework}).
This new framework incorporates three different methods from the literature:
\begin{itemize}
	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2001) \cite{tessendorfSimulating2001} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
	\item The estimation of the forces applied by the fluid on a solid mesh, mixing together different techniques \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014}.
	\item The simulation of the alteration of a free surface by a solid, inspired by Cords and Staadt works (2009) \cite{cordsRealTimeOpenWater2009}.
\end{itemize}
To bring together these previous works, this new framework proposes two new contributions:
\begin{itemize}
	\item An algorithm to compute the ocean velocity at any depth from the simulation of the free surface. This velocity is applied to the fluid to solid forces.
	\item An algorithm to compute the input of the Cords and Staadt method more realistically from the free surface.
\end{itemize}

@@

\subsection{Ocean free surface simulation}
\label{subsubsec:oceanFreeSurfaceSim}
J. Tessendorf (2001) \cite{tessendorfSimulating2001} presented what could be considered a seminal work in real-time ocean simulation. His work relies on the Fast Fourier Transform (FFT) and statistical wave spectra to accurately find the analytical solution to an approximation of the Navier-Stokes equations on a 2D height field. This method is commonly known as the “Tessendorf method”.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solid to fluid coupling}
\label{subsubsec:solidToFluidCoupling}
The simulation of the surface of water in contact with a solid body has been an active research topic for the last twenty years.
Gomez (2000) \cite{gomezInteractiveSimulationWater2000} solved the wave equation using the finite-difference method (FDM) to simulate waves on the free surface.
This method was improved by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to work with moving objects.
Tessendorf (2004) \cite{tessendorfInteractiveWaterSurface2004} proposed a convolutional method.
He extended his method by using an exponential solver \cite{tessendorfEWaveUsingExponential2014}.
Yuksel (2007) \cite{yuksel2007} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
Chentanez and Müller (2010) \cite{chentanezRealTimeSimulationLarge2010} created a hybrid solver combining Eulerian and Lagrangian approaches.
Canabal  \etal (2016) \cite{canabalDispersionKernelsWater2016} improved the physical accuracy of the convolutional method by introducing a dispersion kernel.
Jeschke and Woktan (2017) \cite{jeschkeWaterWavePackets2017} took the work of C. Yuskel a bit further by introducing the notion of a wave packet, which carries information about an entire wave train.
One year later, Jeschke  \etal (2018) \cite{jeschkeWaterSurfaceWavelets2018} proposed to use a wavelet transformation that discretizes the wave amplitudes as a function of space, frequency and direction.
Finally, Schreck  \etal (2019) \cite{schreckFundamentalSolutionsWater2019} based their work on the method of fundamental solutions to generate ambient waves on a large scale domain interacting with complex boundaries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Water height}
\label{subsec:waterHeight}
Tessendorf (2001) \cite{tessendorfSimulating2001} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
\begin{equation}
	\label{equa:heightWater}
	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(\mathbf{k}, t) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
\end{equation}
where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
\begin{equation}
	\label{equa:heightWaterBis}
	\tilde{h}(\mathbf{k}, t) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
\end{equation}
where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see Appendix~\ref{sec:annexe} for more details.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many different \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
In its original paper, Tessendorf (2001) \cite{tessendorfSimulating2001} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
\begin{equation}
	\tilde{h}_0(\mathbf{k}) = \xi_{\mathbf{k}}\sqrt{\frac{4\pi}{L_ik}S(\omega)D(\omega, \theta)\left|\frac{\partial\omega(k)}{\partial k}\right|}
\end{equation}
where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphes \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.

@@

\Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x\ D_z\right]^T$ of seawater to simulate more realistic choppy waves.
As proposed in Sections 4.5 and 4.6 of Tessendorf (2001) \cite{tessendorfSimulating2001}, it is calculated as follows:
\begin{equation}\label{equa:hori-displacement}
	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \frac{i\mathbf{k}}{k} \tilde{h}(\mathbf{k}, t)\exp\left(i\mathbf{k}\cdot\mathbf x\right).
\end{equation}
